Curvature approximation on surfaces -- A Discrete Exterior Calculus Approach

### 2014-12-11 article/main.tex
...

    The equations \eqref{eqDisWeingarten} can be applied, only if we know the exact normals of the manifolds on the vertices, 
    e.g. if the surface \( M \) is described by the 0-level-set of a signed distance function \( \varphi:\R^{3}\rightarrow\R \),
    then is
    \begin{align}
      \vec{\nu} &= \frac{\nabla_{\R^{3}}\varphi}{\left\| \nabla_{\R^{3}}\varphi \right\|} \formpunkt
    \end{align}
    For arbitrary surface triangulations, where the smooth manifold is unknown, like in figure \ref{figBunnyHorseCurvature},
    we must approximate the normals on the vertices before.
    A simple way to do that, is to average the element normals \( \vec{\nu}^{\sigma^{2}} \) of a triangle \( \sigma^{2}\in K \)
    over the Voronoi area:
    \begin{align}
      \left\langle *\vec{\nu}_{\avn} , \star v \right\rangle
        &:= \sum_{\sigma^{2}\succ v} \left| \star v \cap \sigma^{2} \right| \vec{\nu}^{\sigma^{2}} \formpunkt
    \end{align}
    With the same procedure above, we get the approximated curvatures \( K_{W,\avn} \) and \( H_{W,\avn} \).

@@

  \begin{figure}
    \centering\includegraphics[width=.49\textwidth]{bilder/ellipsoid/ErrView.jpg}
    \caption{Ellipsoid \eqref{eqEllipsoid}:
                  The maximal errors for \( K_{W,\avn} \) keeps primarily  on the lateral 4 of all 12 vertices with 1-rings of 5 triangles.
                  100002 DOFs were used (\( h\approx 0.026 \)).
                  The errors for \( K_{GB} \) and \( H_{W,\avn} \) results in a similar picture.}
    \label{figEllipsoidErrView}
  \end{figure}

  \begin{figure}
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/K.jpg}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/H.jpg}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/L2K.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/L2H.eps}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/LMaxK.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/LMaxH.eps}
    \end{minipage}
    \caption{Quartic Surface \eqref{eqQuartic}: Top row: Gaussian curvature \( K \) (left) and mean curvature \( H \) (right).
                              Middle and bottom row: Relative \( L_{2} \) resp. \( L_{\infty} \) error for \( K \) (left) and
                                                     \( H \) (right) for different \( h \) in a log-log-plot.}
    \label{figQuartic}
  \end{figure}

  \begin{figure}
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/K.jpg}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/H.jpg}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/L2K.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/L2H.eps}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/LMaxK.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/LMaxH.eps}
    \end{minipage}
    \caption{Torus: Top row: Gaussian curvature \( K \) (left) and mean curvature \( H \) (right).
                              Middle and bottom row: Relative \( L_{2} \) resp. \( L_{\infty} \) error for \( K \) (left) and
                                                     \( H \) (right) for different \( h \) in a log-log-plot.}
    \label{figTorus}
  \end{figure}

>>>>

    The equations \eqref{eqDisWeingarten} can be applied, only if we know the exact normals of the manifolds on the vertices, 
    e.g. if the surface \( M \) is described by the 0-level-set of a signed distance function \( \varphi:\R^{3}\rightarrow\R \),
    then is
    \begin{align}
      \vec{\nu} &= \frac{\nabla_{\R^{3}}\varphi}{\left\| \nabla_{\R^{3}}\varphi \right\|} \formpunkt
    \end{align}
    For arbitrary surface triangulations, where the smooth manifold is unknown, like in figure \ref{figBunnyHorseCurvature},
    we must approximate the normals on the vertices before.
    A simple way to do that, is to average the element normals \( \vec{\nu}^{\sigma^{2}} \) of a triangle \( \sigma^{2}\in K \)
    over the Voronoi area:
    \begin{align}
      \left\langle *\vec{\nu}_{\avn} , \star v \right\rangle
        &:= \sum_{\sigma^{2}\succ v} \left| \star v \cap \sigma^{2} \right| \vec{\nu}^{\sigma^{2}} \formpunkt
    \end{align}
    With these discrete normals on the vertices and the same procedure above, 
    we get then the approximated curvatures \( K_{W,\avn} \) and \( H_{W,\avn} \).

  \subsection{Gauss-Bonnet-Operator}
    Thus far, we have two approaches for the mean curvature with sections \ref{secCurvatureVector} and \ref{secWeingarten}.
    For a second approach of the Gauss curvature, we want to use a Gauss-Bonnet approximation, like in \cite{polthier} or \cite{octavian}, for comparison reasons.

    On the abstract Voronoi cell \( \pi(\star v) \) (arise by projecting to the manifold) of a vertex \( v \), 
    it holds the Gauss-Bonnet theorem \cite{berger}:
    \begin{align}
      \label{eqGaussBonnet}
      \int_{\pi(\star v)}K \mu = 2\pi - \sum_{i=1}^{m_{v}}\tilde{\beta_{i}} - \int_{\partial\pi(\star v)} k_{g} ds \formpunkt
    \end{align}
    \( \mu\in\Omega^{2}(M)  \) is the volume form, \( k_{g} \) the geodetic curvature on the boundary edges of the cell 
    and \( \tilde{\beta_{i}} \) the \( m_{v} \) exterior angles of the boundary on the Voronoi vertices.
    This can be approximated on the simplicial complex \( K \) by omitting the projection.
    The geodetic curvature \( k_{g} \) vanish on the straight lines and on the bucklings, where the dual edges intersect the primal edges.
    Hence, in our DEC notation, we get
    \begin{align}
      \label{eqDisGaussBonnet}
      \left\langle *K_{GB} , \star v \right\rangle &:= 2\pi - \sum_{i=1}^{m_{v}}\beta_{i}
    \end{align}
    on a vertex \( v\in K \), with the exterior angles \( \beta_{i} \).
    \begin{figure}
      \begin{minipage}[htp]{.235\textwidth}
        \centering\includegraphics[width=.99\textwidth]{bilder/tikz/GaussBonnet.eps}
      \end{minipage}
      \hfill
      \begin{minipage}[htp]{.235\textwidth}
        \centering\includegraphics[width=.99\textwidth]{bilder/tikz/ElementGaussBonnet.eps}
      \end{minipage}
      \caption{Left: \(  m = m_{v} =: \tilde{m}+1 \) exterior angles \( \beta_{i} \) on the vertices of a Voronoi cell.
                The solid lines (with orientations) marks the dual (Voronoi) edges.
               Right: Angles in a primal triangle referring to the edges from the circumcenter subdivision.}
      \label{figGaussBonnet}
    \end{figure}
    We see in figure \ref{figGaussBonnet}, that it is possible to calculate the exterior angles by the local interior angles 
    \( \alpha_{0} \) and \( \alpha_{1} \) on \( v \) for a
    triangle \( \sigma^{2}\succ v \):
    \begin{align}
      \beta_{\bullet} = \pi - \gamma_{0} - \gamma_{1} = \alpha_{0} + \alpha_{1} \formpunkt
    \end{align}
    Because the interior angles are determinable by the length measures of the edges, we get 
    \begin{align}
      \sum_{i=1}^{m_{v}}\beta_{i} &= \sum_{\sigma^{2}\succ v} \sum_{\sigma^{2}\succ\sigma^{1}\succ v}
                               \text{atan2}\left( 2\left|\star\sigma^{1} \cap \sigma^{2} \right|, 
                                 \left|\sigma^{1}\right|\right)
    \end{align}
    for the angle sum in \eqref{eqDisGaussBonnet}.

@@

  \begin{figure}
    \centering\includegraphics[width=.49\textwidth]{bilder/ellipsoid/ErrView.jpg}
    \centering\includegraphics[width=.49\textwidth]{bilder/ellipsoid/ErrViewHLX.jpg}
    \caption{Ellipsoid \eqref{eqEllipsoid}:
                  The maximal errors for \( K_{W,\avn} \) (top) keeps primarily  on the lateral 4 of all 12 vertices with 1-rings of 5 triangles.
                  The errors for \( K_{GB} \) and \( H_{W,\avn} \) results in a similar picture.
                  Slightly different is the situation at the errors for \( H_{LX} \) (bottom).
                  There are the maximums partially on vertices in the neighbourhood. 
                  100002 DOFs were used (\( h\approx 0.026 \)).}
    \label{figEllipsoidErrView}
  \end{figure}

  \begin{figure}
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/K.jpg}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/H.jpg}
    \end{minipage}
    \caption{Quartic surface\eqref{eqQuartic}: Gaussian curvature \( K \) (left) and mean curvature \( H \) (right).}
    \label{figQuartic}
  \end{figure}

  \begin{figure}
    \centering\includegraphics[width=.49\textwidth]{bilder/quartic/L2K.eps}
    \centering\includegraphics[width=.49\textwidth]{bilder/quartic/L2H.eps}
    \caption{Quartic surface \eqref{eqQuartic}: Relative \( L_{2} \) error for \( K \) (top) and
                                                     \( H \) (bottom) for different \( h \) in a log-log-plot.}
    \label{figQuarticL2Err}
  \end{figure}

  \begin{figure}
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/K.jpg}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/H.jpg}
    \end{minipage}
    \caption{Torus: Gaussian curvature \( K \) (left) and mean curvature \( H \) (right).}
    \label{figTorus}
  \end{figure}

  \begin{figure}
    \centering\includegraphics[width=.49\textwidth]{bilder/torus/L2K.eps}
    \centering\includegraphics[width=.49\textwidth]{bilder/torus/L2H.eps}
    \caption{Torus: Relative \( L_{2} \) error for \( K \) (top) and
                                                     \( H \) (bottom) for different \( h \) in a log-log-plot.}
    \label{figTorusL2Err}
  \end{figure}
